### 2020-11-28 theorems.tex

\newcommand\PartOfBoundOffDiagonal{3\efourB+4(K + \dmax)\,\efive}
\newcommand\boundOffDiagonal{\efourA + \PartOfBoundOffDiagonal}
\newcommand\K{K^\prime}
\newcommand\ThmOneVarBound{Q^{4b} \, \Unit\Time  \big(\altb + \etwo \, (\Unit\Time-\altb) \big)}

\newcommand\BvConstant{\mathcal{C}_0}
\newcommand\dmax{d_{\max}}
\newcommand\boundLemma{\efourB+(K+d_{\unit,\time})\efive}
\newcommand\boundLemmaUniform{\efourB+(K+\dmax)\efive}

%\newcommand\altB{\tilde{B}}
\newcommand\altB{C}
%\newcommand\el{\epsilon_\ell}
\newcommand\el{\epsilon}   %%% epsilon for Thm 1
%\newcommand\eone{\epsilon_1}
%\newcommand\eone{\epsilon^{}_{\!\mathrm{\ref{A1}}}}
\newcommand\eone{\epsilon^{}_{\!\mathrm{A1}}}
%\newcommand\eone{\epsilon^{A}_{1}}
%\newcommand\etwo{\epsilon^{}_{\!\mathrm{\ref{A:unconditional:mix}}}}
\newcommand\etwo{\epsilon^{}_{\!\mathrm{A4}}}
%\newcommand\altb{\tilde{b}}
\newcommand\altb{c}
\newcommand\Ai{
\begin{assumption}\label{A1}
There is an $\eone>0$, independent of $\Unit$ and $\Time$, and a collection of neighborhoods $\{B_{\unit\comma\time}\subset A_{\unit\comma\time}, \unit\in 1\mycolon\Unit,\time\in 1\mycolon\Time\}$ such that, for all  $\unit$ and $\time$, any  bounded real-valued function $|h(x)|\le 1$, and any value of $x_{B^c_{\unit\comma\time}}$, 
%\begin{equation}
\begin{eqnarray}
\nonumber
&& \hspace{-12mm} \Bigg| \int h(x_{\unit\comma\time}) f_{X_{\unit\comma\time}|Y_{B_{\unit\comma\time}},X_{B^c_{\unit\comma\time}}}(x_{\unit\comma\time}\given  \data{y}_{B_{\unit\comma\time}}, x_{B^c_{\unit\comma\time}}) \, dx_{\unit\comma\time}
\\
\siOnly{\label{eq:weak_coupling2}}
\msOnly{\nonumber}
&& \hspace{-3mm} - \int h(x_{\unit\comma\time})f_{X_{\unit\comma\time}|Y_{B_{\unit\comma\time}}}(x_{\unit\comma\time}\given \data{y}_{B_{\unit\comma\time}})
\, dx_{\unit\comma\time}
\Bigg| \hspace{1mm} < \hspace{1mm} \eone.
%\end{equation}
\end{eqnarray}
\end{assumption}
}

\newcommand\Aii{
\begin{assumption}\label{A1b}
For the collection of neighborhoods in Assumption~\ref{A1}, with $B^+_{\unit\comma\time}=B^{}_{\unit\comma\time}\cup (\unit,\time)$, there is a constant $\Bsize$, depending on $\eone$ but not on $\Unit$ and $\Time$, such that
\begin{equation}
\nonumber
\sup_{\unit\in 1:\Unit, \, \time\in 1:\Time} \big| B^+_{\unit\comma\time}\big| \le \Bsize.
\end{equation}
\end{assumption}
}

\newcommand\Aiii{
\begin{assumption}\label{A2}
There is a constant $Q$, independent of $\Unit$ and $\Time$, such that, for all $\unit$ and $\time$,
\begin{equation}
\nonumber
Q^{-1} <f_{Y_{\unit\comma\time}|X_{\unit\comma\time}}(\data{y}_{\unit\comma\time}\given x_{\unit\comma\time}) < Q
\end{equation}
\end{assumption}
}

\newcommand\Aiv{
  \begin{assumption}\label{A:unconditional:mix}
    There exists $\etwo > 0$, independent of $\Unit$ and $\Time$, such that the following holds.
    For each $\unit, \time$, a set $\altB^{}_{\unit\comma\time} \subset (1\mycolon U)\times(0\mycolon N)$ exists such that $(\altUnit, \altTime) \notin \altB_{\unit\comma\time}$ implies $B_{\unit\comma\time}^+ \cap B_{\altUnit\comma\altTime}^+ = \emptyset$ and
      \begin{equation} 
\siOnly{\label{eq:A:unconditional:mix}}
\msOnly{\nonumber}
        \big| f_{X_{B^{+}_{\altUnit\comma\altTime}}|X^{}_{B^{+}_{\unit\comma\time}}} - f_{X_{B^{+}_{\altUnit\comma\altTime}}} \big|
        < \etwo \, f_{X_{B^{+}_{\altUnit\comma\altTime}}}
      \end{equation}
      Further, there is a uniform bound $|\altB_{\unit\comma\time}| \le \altb$.
\end{assumption}
}

\newcommand\TheoremI{
\begin{theorem}  \label{thm:tif}
Let $\MC{\loglik}$ denote the Monte Carlo likelihood approximation constructed by {\UBF}.
Consider a limit with a growing number of bootstrap replicates, $\Rep\to\infty$, and suppose assumptions~\ref{A1},~\ref{A1b} and~\ref{A2}.
There are quantities $\el(\Unit,\Time)$ and $V(\Unit,\Time)$, with bounds $|\el|<\eone Q^{2}$ and $V < Q^{4\Bsize} \, U^2 N^2$, such that
\begin{equation}
\label{th1:lik:bound}
\Rep^{1/2}\big[ \MC{\loglik}-\loglik-\el\Unit\Time \big]  \xrightarrow[\Rep \rightarrow \infty]{d} \normal\big[0,V\big],
\end{equation}
where $\xrightarrow[\Rep \rightarrow \infty]{d}$ denotes convergence in distribution and $\normal[\mu,\Sigma]$ is the normal distribution with mean $\mu$ and variance $\Sigma$.
If additionally Assumption~\ref{A:unconditional:mix} holds, we obtain an improved variance bound
\begin{equation}
\label{th1:lik:bound2}
V < \ThmOneVarBound.
\end{equation}
\end{theorem}
}

%%% 2222222222222222222222

%\newcommand\ethree{\epsilon^{}_{\mathrm{\ref{B1}}}}
\newcommand\ethree{\epsilon^{}_{\mathrm{B1}}}
%\newcommand\efourA{\epsilon^{}_{\mathrm{\ref{B:unconditional:mix}}}}
\newcommand\efourA{\epsilon^{}_{\mathrm{B4}}}
%\newcommand\efourB{\epsilon^{}_{\mathrm{\ref{B:temporal:mix}}}}
\newcommand\efourB{\epsilon^{}_{\mathrm{B5}}}
%\newcommand\efive{\epsilon^{}_{\mathrm{\ref{B:girf}}}}
\newcommand\efive{\epsilon_{\mathrm{B6}}}

\newcommand\Bi{

\begin{assumptionB}\label{B1}
  There is an $\ethree>0$, independent of $\Unit$ and $\Time$, and a collection of neighborhoods $\{B_{\unit\comma\time}\subset A_{\unit\comma\time}, \unit\in 1\mycolon\Unit,\time\in 1\mycolon\Time\}$ such that the following holds for all $\unit$ and $\time$, and any bounded real-valued function $|h(x)|\le 1$\emph{:} if we write $A=A_{u,n}$, $B=B_{u,n}$, $f_{A}(x_A) = f_{Y_{A}|X_{A}}(\data y_{A}|x_A)$, and $f_{B}(x_B) = f_{Y_{B}|X_{B}}(\data y_{B}|x_B)$,
%\begin{equation}
\begin{eqnarray}
\nonumber
%\label{asif:eq:weak_coupling2}
&&\hspace{-7mm}
\left| 
\int \hspace{-1mm} h(x)
\left\{
\frac{
  \E_{g} \! \big[ 
    f_{A}\big(X_{A}^P\big) f_{X_{u,n}|X_{A^{[n]}},\myvec X_{n-1}} \big(x | X_{A^{[n]}}^P, \myvec X_{n-1}\big) 
  \big]
} {
  \E_{g} \! \big[f_A\big(X_A^P\big)\big]
} - 
\right.
\right.
\\
\nonumber
&& 
\siOnly{\hspace{20mm}}
\hspace{-6.5mm} 
% - \hspace{-1mm} 
\siOnly{\hspace{2mm}}
\left.
\left.
\frac{
  \E_{g} \! \big[
    f_{B}\big(X_{B}^P\big) f_{X_{u,n}|X_{B^{[n]}},\myvec X_{n-1}}\big(x | X_{B^{[n]}}^P, \myvec X_{n-1}\big) \big] 
} {
  \E_{g} \! \big[f_B\big(X_B^P\big)\big]
} 
\!
\right\}
\!
dx
\right| \hspace{0mm} < \hspace{0mm} \ethree.
%\end{equation}
\end{eqnarray}
\end{assumptionB}
}

\newcommand\Bii{
\begin{assumptionB}\label{B1b}
The bound $\sup_{\unit\in 1:\Unit,\time\in 1:\Time} \big| B^+_{\unit\comma\time}\big| \le \Bsize$ in Assumption~\ref{A1b} applies for the neighborhoods defined in Assumption~\ref{B1}. 
This also implies there is a finite maximum temporal depth for the collection of neighborhoods, defined as
\begin{equation}
\nonumber
\dmax=\sup_{(\unit,\time)}\hspace{1mm} \sup_{(\altUnit,\altTime)\in B_{\unit,\time}}|\time-\altTime|.
\end{equation}
\end{assumptionB}
}

\newcommand\Biii{
\begin{assumptionB}\label{B2} Identically to Assumption~\ref{A2}, 
$Q^{-1} <f_{Y_{\unit\comma\time}|X_{\unit\comma\time}}(\data{y}_{\unit\comma\time}\given x_{\unit\comma\time}) < Q$.
\end{assumptionB}
}

\newcommand\BivA{
\begin{assumptionB}\label{B:unconditional:mix}
Define $g_{{X}_{A}|{X}_{B}}$ via \myeqref{eq:g}.
Suppose there is an $\efourA$, independent of $\Unit$ and $\Time$, such that the following holds.
For each $\unit$ and $\time$, a set $\altB^{}_{\unit\comma\time} \subset (1\mycolon U)\times(0\mycolon N)$ exists such that $(\altUnit, \altTime) \notin \altB_{\unit\comma\time}$ implies $B_{\unit\comma\time}^+ \cap B_{\altUnit\comma\altTime}^+ = \emptyset$ and
%      \begin{eqnarray} 
%\label{eq:B:unconditional:mix:1}
\begin{equation}
\nonumber
\big| 
   g^{}_{{X}^{P}_{B^{}_{\altUnit\comma\altTime}}{X}^{P}_{B^{}_{\unit\comma\time}}} 
   - g^{}_{{X}^{P}_{B^{}_{\altUnit\comma\altTime}}} \, g^{}_{{X}^{P}_{B^{}_{\unit\comma\time}}} 
\big|
<
%&\hspace{-3mm}<\hspace{-3mm} & 
%\\
%&&\hspace{-30mm}
(1/2) \, \efourA \, \,  g^{}_{{X}^{P}_{B^{}_{\altUnit\comma\altTime}}} \, g^{}_{{X}^{P}_{B^{}_{\unit\comma\time}}}
%\\
%\label{eq:B:unconditional:mix:2}
\end{equation}
\begin{eqnarray}
\nonumber
\big|
g^{}_{X^P_{B^{}_{\altUnit\comma\altTime}}|\myvec{X}_{0:\Time}} 
\, g^{}_{X^P_{B^{}_{\unit\comma\time}}|\myvec{X}_{0:\Time}} 
-
g^{}_{X^P_{B^{}_{\altUnit\comma\altTime}},X^P_{B^{}_{\unit\comma\time}}|\myvec{X}_{0:\Time}}
\big|
&&
%&\hspace{-3mm}<\hspace{-3mm}& 
\\
\nonumber
%&&\hspace{-30mm}(1/2) \, \efourA \, \, 
&&\hspace{-30mm} < (1/2) \, \efourA \, \, 
 g^{}_{X^P_{B^{}_{\altUnit\comma\altTime}},X^P_{B^{}_{\unit\comma\time}}|\myvec{X}_{0:\Time}}
      \end{eqnarray}
      Further, there is a uniform bound $|\altB_{\unit\comma\time}| \le \altb$.
\end{assumptionB}
}
\newcommand\BivB{
\begin{assumptionB}\label{B:temporal:mix}
There is a constant $K$, independent of $\Unit$ and $\Time$, such that, for any $0\leq d \leq \dmax$, any $\time\ge K+d$, and any set $D\subset (\UnitSet)\times(n{\mycolon}n-d)$, 
\begin{eqnarray}
\nonumber
&&\hspace{-6mm} \big|
g^{}_{{X}_{D}|\myvec{X}_{\time-d-K}}(x^{}_{D}\given\myvec{x}^{(1)}_{\time-d-K})
-
g^{}_{{X}_{D}|\myvec{X}_{\time-d-K}}(x^{}_{D}\given\myvec{x}^{(2)}_{\time-d-K})
\big|
\\
\nonumber
&& \siOnly{\hspace{10mm}} \hspace{4mm}
< 
\efourB  \,  \,
g^{}_{{X}_{D}|\myvec{X}_{\time-d-K}}(x^{}_{D}\given\myvec{x}^{(1)}_{\time-d-K})
\end{eqnarray}
holds for all $\myvec{x}^{(1)}_{\time-d-K}$, $\myvec{x}^{(2)}_{\time-d-K}$, and $x^{}_{D}$.
\end{assumptionB}
}

\newcommand\Bv{
\begin{assumptionB}\label{B:girf}
Let $h$ be a bounded function with $|h(x)|\le 1$. 
Let $\myvec{X}^{\GR}_{\time,\Ninter,\np,\rep}$ be the Monte Carlo quantity constructed in {\ABFIR}, conditional on $\myvec{X}^{\IF}_{\time-1,\Ninter,\rep}=\myvec{x}^{\IF}_{\time-1,\Ninter,\rep}$.
There is a constant $\BvConstant(\Unit,\Time,\Ninter)$ such that, for all $\efive^{}>0$ and $\myvec{x}^{\IF}_{\time-1,\Ninter,\rep}$, whenever the number of particles satisfies $J > \BvConstant(\Unit,\Time,\Ninter)/\efive^{3}$,
\begin{equation}
\nonumber
\hspace{-0.5mm}
\left|  
\, 
  \EMC\Big[ 
    \frac{1}{\Np} \sum_{\np=1}^{\Np} h(\myvec{X}^{\GR}_{\time,\Ninter,\np,\rep}) 
   \Big] \!
-  \E_g \! \big[h(\myvec{X}_\time)\given \myvec{X}_{\time-1}=\myvec{x}^{\IF}_{\time-1,\Ninter,\rep}\big] 
\right| \!
< \efive^{}.
\end{equation}
\end{assumptionB}
}

\newcommand\Bvi{
\begin{assumptionB}\label{B:XG_XA_ind}
For $1\leq \time \leq N$, the Monte Carlo random variable $X^A_{\time,i}$ is independent of $w^M_{\unit,\time,i,j}$ conditional on $X^A_{\time-1,i}$.
\end{assumptionB}
}

\newcommand\TheoremII{
\begin{theorem}  \label{thm:asif}
Let $\MC{\loglik}$ denote the Monte Carlo likelihood approximation constructed by {\ABFIR}, or by {\ABF} since this is the special case of {\ABFIR} with $\Ninter=1$.
Consider a limit with a growing number of bootstrap replicates, $\Rep\to\infty$, and suppose assumptions~\ref{B1},~\ref{B1b}, \ref{B2}, \ref{B:temporal:mix}, \ref{B:girf} and~\ref{B:XG_XA_ind}. 
Suppose the number of particles $\Np$ exceeds the requirement for~\ref{B:girf}.
There are quantities $\el(\Unit,\Time)$ and $V(\Unit,\Time)$ with $|\el|<
Q^2\ethree + 2 Q^{2\Bsize}\big(\boundLemmaUniform\big)$
and $V<Q^{4\Bsize}\Unit^2\Time^2$ 
such that
\begin{equation}
\siOnly{\label{th:asif:lik:bound}}
\msOnly{\nonumber}
\Rep^{1/2}
  \big[ 
    \MC{\loglik}-\loglik-\el\Unit\Time 
  \big]  
  \xrightarrow[\Rep \rightarrow \infty]{d} 
  \normal\big[ 0,V \big].
\end{equation}
If additionally Assumption~\ref{B:unconditional:mix} holds, we obtain an improved rate of
\begin{equation}
\siOnly{\label{th:asif:lik:bound2}}
\msOnly{\nonumber}
V < Q^{4\Bsize}\Time\Unit
\big\{ c + \big(\boundOffDiagonal \big) \big(\Time\Unit-c\big)
\big\}
\end{equation}
\end{theorem}
}